---
title: "How to use BARTSobol: Sobol' indices vs variable counts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r message=FALSE, warning=FALSE}
# devtools::install("/Users/akira/Dropbox/Rpkgs/BARTSobol")
# install.packages("BART")
```


# Overview

This vignette illustrates how to use the BARTSobol package by showcasing the 
performance difference between 
the BART-based Sobol' sensitivity indices calculations found in (XX: cite my paper) 
and variable counts.


# Comparison

Here we compare normalized main-effects and total-effects Sobol' sensitivity 
indices to variable counts of a trained BART model. 
This BART model is trained on data generated from the Friedman function 
\[
f(\mathbf{x}) = \sin(\pi x_1 x_2) + 20(x_3 - 0.5)^2 + 10x_4 + 5x_5
\]
applied to the $2^{12}$-Sobol' sequence on the unit hypercube $[0, 1]^p$ 
where $p = 7$. 
This BART model uses 200 trees and 1000 posterior samples.

The main-effects and total-effects Sobol' sensitivity indices of the Friedman function are:

- $(S_1, S_2, S_3, S_4, S_5, S_6, S_7) = (.197, .197, .093, .350, .087, 0, 0)$.
- $(T_1, T_2, T_3, T_4, T_5, T_6, T_7) = (.274, .274, .093, .350, .087, 0, 0)$.

These indices are shown in grey in the plot below.

The main-effects and total-effects Sobol' sensitivity indices of the 
1,000 posterior samples of the BART model are shown below in black. 


```{r setup}
library(BART)
library(BARTSobol)
library(tidyverse)
```

```{r}
p <- 7
n_obs <- 2^12
n_tree <- 200L
nd_post <- 1000L

# Create training data
x_train <- SobolSequence::sobolSequence.points(dimR = p, count = n_obs)
Fried <- function(x) 10*sin(pi*x[1]*x[2]) + 20*(x[3]-0.5)^2 + 10*x[4] + 5*x[5]
y_train <- apply(x_train, 1, Fried) + rnorm(n_obs)

# Train BART model; call wbartSobol()
wb_obj <- wbart(x_train, y_train, ntree = n_tree, ndpost = nd_post, keepevery = 100L, nskip = 1000L)
wb_sobol <- wbartSobol(wb_obj, p, n_tree, nd_post)

# Plot results
my_titles <- paste("X", 1:p)
cntnorm <- apply(wb_obj$varcount, 1, function(x) x/sum(x))  # apply transposes its argument...
Si <- data.frame(Variable = my_titles, wb_sobol$Si[, -1])
Ti <- data.frame(Variable = my_titles, wb_sobol$TSi[, -1])
cntnorm <- data.frame(Variable = my_titles, cntnorm)
rownames(cntnorm) <- NULL
# ints <- data.frame(Variable = my_titles, wb_sobol$TSi[, -1] - wb_sobol$Si[, -1])

Si_longer <- as.data.frame(Si) %>%
  pivot_longer(-Variable, names_to = "colname", values_to = "value") %>%
  mutate(Metric = "Si") %>% 
  mutate(Model = "BART")
Ti_longer <- as.data.frame(Ti) %>%
  pivot_longer(-Variable, names_to = "colname", values_to = "value") %>%
  mutate(Metric = "Ti") %>% 
  mutate(Model = "BART")
cntnorm_longer <- as.data.frame(cntnorm) %>%
  pivot_longer(-Variable, names_to = "colname", values_to = "value") %>%
  mutate(Metric = "Count") %>% 
  mutate(Model = "BART")
# ints_longer <- as.data.frame(ints) %>%
#   pivot_longer(-Variable, names_to = "colname", values_to = "value") %>%
#   mutate(Metric = "Ti - Si") %>% 
#   mutate(Model = "BART")


# Sobol' index values of the Friedman function 
Si_f <- c(.197, .197, .093, .350, .087, rep(0, times = p-5))
Ti_f <- c(.274, .274, .093, .350, .087, rep(0, times = p-5))
Si_longer <- rbind(Si_longer, 
                   tibble(Variable = my_titles, colname = "f", 
                          value = Si_f, Metric = "Si", Model = "f"))
Ti_longer <- rbind(Ti_longer, 
                   tibble(Variable = my_titles, colname = "f", 
                          value = Ti_f, Metric = "Ti", Model = "f"))
# ints_longer <- rbind(ints_longer,
#                      tibble(Variable = my_titles, colname = "f", 
#                             value = Ti_f - Si_f, Metric = "Ti - Si", Model = "f"))




p <- rbind(Si_longer, Ti_longer, cntnorm_longer) %>%
  ggplot(aes(x = Variable, y = value, shape = Metric, color = Model)) +
  geom_boxplot() +
  theme_bw() +
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) +
  facet_grid(Metric ~ .) + 
  scale_colour_grey()
p
```






# Citation

XX: Include citations

Cite my paper, Sobol' 1993, and CGM 2010. 












